---
title: "Homicide Report USA, 1980-2014"
author: "My Tran, Alex Park, Esmay Muniz"
date: "2022-11"
output: html_document
---

```{r setup, include=FALSE,message=F}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
library(tidyverse)
```

# Introduction


[Link here for the murder data set](https://www.kaggle.com/datasets/murderaccountability/homicide-reports)


When reading about crime incidents or watching them, do you guys ever think about the statistics of how much more men are prone to kill than women, or why is that killers choose a particular weapon to kill or ever thought about the correlations where a perpetrator and a victim know each-other? In the data that we decided to work with is [US homicide reports 1980-2014]. This dataset includes murders from the FBI's Supplementary Homicide Report from 1976 to the present and Freedom of Information Act data. Some of this data some cases were either solved or unsolved. It also includes homicides that were not reported to the Justice Department.

This dataset was the most interesting and the most current data available in which contain victims' age, perpetrators age, location, state, year, weapon used, and data that align with our interest of what factors correlate with behaviors and motives of the killers. Our approach is to analyze the data and filter the dataset in order to answer the following questions; at what rate has the homicide data change since 1980?, what factors contribute to the year with the most kills? and what motives and behaviors contribute to a killer attacking their victim's. Through our analyzes we will find approaches to our findings. Our data will give a visualization and identify correlations within these murders throughout the year of 1980-2014.



# Our Questions 

__Domain Question__

What factors are related to __motives__ and __behaviors__ of the killers? 

__Other questions__

What __weapon__ was used the most?

What __state__ had the most kills?

Do the victims __know__ their perpetrator?

How does __gender__ play a role in homicide incidents?


# Get ready

```{r, message = F, warning=F}
library(tibble) # used to create tibbles
library(tidyr) # used to tidy up data
library(rmarkdown) # dynamic document 
library(ggplot2) # used for data visualization
library(dplyr) # used for data manipulation
library(shiny) # used for showing dynamic visuals in collaboration with ggvis
library(prettydoc)# used for creating pretty documents from R markdown
library(knitr)#for dynamic report generation
library(tidyverse)# multiple tidy up data packages here
library(hms) # used to install kableExtra package
library(kableExtra) # used to construct Complex Table for data
library(dplyr) # used to install tigris package
library(tigris) # used to make states map
#added library for other graphs
library(plotly)
library(rjson)
library(leaflet)
library(leaflet.providers)
library(maps)
library(viridis)
library(viridisLite)
library(sp)
library(quantmod)
library(plot3D)
library(sf)
library(RColorBrewer)
library(gganimate)
```


# Unfiltered data


__Original dataset__ we have is from Kaggle, "Homicide Report". Firstly, we had to download our data from Kaggle, which came in a zip file. Once, when we created an R Markdown file and we saved it into a new folder. We extract the zip into the new folder in order to obtain the CSV (comma-separated values file) called database.csv, which holds the data.

```{r}
unzip(zipfile="./homicide.zip")
data <- read.csv("database.csv")
glimpse(data) 
```


## Data Key Terms:

__Agency Type__: Law enforcement Agency who handled the case

__State/City__: State and Counties of the reported homicides

__Year/Month__: Time stamp of the homicides

__Crime Type__: Murder, Manslaughter or Negligence designated to case

__Crime Solved__: Whether the case has been solved or not

__Victim Sex/Age/Race__: Victim profile

__Perpetrator Sex/Age/Race__: Perpetrator profile

__Relationship__: The perpetrators relation to the victim

__Weapon__: Weapon used to commit homicide

__Case Open/Closed__: Change the designation of a crime being solved.

__Solve Rate__: Percentage of Homicide Reports where the case was closed



## 1. Top murder cases by state:

__Top murder cases by state,__ we are taking our unfiltered data to display the top murder cases by State to compare which states have the most murders. To do this, we needed to plot our data by grouping by state. We then needed to show the total of murders per state and arranged by murder count from descending order. 

```{r, message=FALSE}
data %>% group_by(State) %>% 
  summarize(Murder_Count = n()) %>% 
  arrange(desc(Murder_Count)) %>% 
  kbl() %>% kable_paper()  %>% scroll_box(height = "300px")
```



```{r, echo=F,fig.height=7, fig.width=9, message=F}

data %>%
  plot_ly(x = ~State) %>% 
  add_histogram(color = I("red"), 
                stroke = I("black"), 
                span = I(2)
                , alpha = 0.6) %>% group_by(State) %>%
  summarise(Murder_Count = n()) %>% 
  layout(title = "Histogram of States count by Murder Cases", xaxis = list(title = "States"), yaxis = list(title = "Murder Case Count"))
```


After we group all the states in the US and the number count of murders, we plot them into a interactive histogram. It helped to see how much more murders were in California than other states. Also Texas and NY, showed the spiked than the other states.

In table and graph, they show us that the state with the most kills in which California is the state with the most murder cases with Texas coming in second and New York third.


## 2. Top 10 States with the highest amount of cases.

```{r echo=FALSE, fig.height=7, fig.width=9, message=FALSE}
                                                                 
top10_Murder_Cases <- data %>% group_by(State) %>% summarize(Murder_Count = n()) %>% arrange(desc(Murder_Count)) %>% head(10) 
top10_Murder_Cases

# I shown the top Murder Cases count in the Histogram to compare 
top10_Murder_Cases %>% plot_ly(x = ~Murder_Count, y = ~State, color = I("dark red"), 
                stroke = I("black"), 
                span = I(2)
                , alpha = 0.6)%>% layout(title = "Histogram of top 10 States By Murder Case", xaxis = list(title = "States"), yaxis = list(title = "Murder Case Count"))

```


We then wanted to focus on the top ten most murder cases per state to help optimize the differences. We made another interactive histogram, you can see California and Texas having the most murder cases. They are one of the largest states in the US, so it helps to know why they would have the most kills. New York is not as a massive state as TX or CA, but it still ranked number three in the most murder cases. Our assumption is since NY is one of the biggest city in the US, the city that never sleeps and one the of the most tourist city, there would be more crimes than those with smaller states. By seeing the histogram, it brought up the question to us wanting to know what weapon is being used in this murder cases ? 


## 3. What is the most used weapon?

We count how many times killers used each kind of weapons to see the top of their weapon choice.

```{r}
data %>% group_by(Weapon) %>% 
  summarize(Most_Weapon_Used = n()) %>% 
  arrange(desc(Most_Weapon_Used)) %>% 
  kbl() %>% kable_paper() %>% scroll_box(height = "300px")
```


Here you can see that the handgun is the most "favorite" weapon of serial killers compare to other weapons. We also have a good amount of "unknown" weapon, we will need to clean out our data to make it more cleaner and get more accurate results.



## 4. Murder Cases in USA

Here we wanted to visualize the highest crime counts in the US. Heatmaps are great when focusing on locations that matter the most. In this case, we see CA being red compare to other states. Also, in this heatmap, it shows how in Northern US there is less crime count. Lasty, if you look at the north side of the US, there are less murder cases compare to other regions. 

```{r echo=FALSE, results=FALSE, fig.width=18, fig.height=15, message=FALSE, warning=F }
states <- states(cb = TRUE, class = "sf") %>% 
  filter(!as.numeric(STATEFP) %in% c(2, 15, 60, 66, 69, 72, 78)) %>%
  mutate(var = rnorm(nrow(.)))
data_murder_by_states <- data %>% group_by(State) %>% 
  summarize(murder = n()) %>% arrange(desc(murder))
heatmap_data_murder_by_states <- states %>% left_join(data_murder_by_states, by = c('NAME' = 'State'))
ggplot(heatmap_data_murder_by_states) + 
  geom_sf(aes(fill = murder ), colour = alpha("black", 1 / 2), size = 0.2) +   
  scale_fill_viridis_c(option = "turbo" , 
                       begin = 0.1, name = "Murder Count",label = scales::comma) + geom_sf_text(aes(label = NAME), colour = "black") + 
  labs(title = "Heat Map of the United States", 
      subtitle = "Count by Murder in each State") + 
  theme(panel.background = element_rect(colour = "black")) + 
theme_bw() 
```

## 5. Murder Cases in Texas
 
Now let's focus on the best state of the US, Texas. Unfortunately, Texas comes in second with the biggest crime rates. We wanted to see what county/city had the biggest crime rate. Kaufman county of Texas had the highest crime rate.

```{r, fig.width=19, fig.height=16, echo=FALSE, results=FALSE, message=FALSE, warning=FALSE, Error=FALSE}


Texas <- tigris:: counties(state = "TX", cb = T, class ='sf')

Data_City_by_Murder <- data %>% group_by(City) %>% summarize(Murder = n()) %>% arrange(desc(Murder))

heat_map_of_Texas <- Texas %>% left_join(Data_City_by_Murder, by = c("NAME" = 'City'))


ggplot(heat_map_of_Texas) + geom_sf(aes(fill = Murder)) + scale_fill_viridis_c(option = "inferno" , begin = 0.1, label = scales::comma) + stat_sf_coordinates() + theme(legend.position="right",
        plot.title = element_text(hjust = 0.5,color = "dark red", size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, color = "dark red", size = 14, face = "bold"),
        plot.caption = element_text(color = "Gray60"))+ 
  guides(fill = guide_legend(title = "Murder Count", title.position = "top", title.theme =element_text(size = 10, face = "bold",colour = "Black",angle = 0))) + geom_sf_text(aes(label = NAME), colour = "grey", alpha = 0.6) + labs(title = "Murder Count by Counties in Texas", subtitle = "1980 to 2014")



```


We wanted to include California since it has the highest among all other states to see where most of the murders are.  
California is broken down into cities instead of counties.

```{r, fig.width=19, fig.height=16, echo=FALSE, results=FALSE, message=FALSE, warning=FALSE, Error=FALSE}


California <- tigris:: counties(state = "CA", cb = T, class ='sf')
Data_City_by_Murder <- data %>% group_by(City) %>% summarize(Murder = n()) %>% arrange(desc(Murder))
heat_map_of_California <- California %>% left_join(Data_City_by_Murder, by = c("NAME" = 'City'))

# This is the heatmap for California 
ggplot(data = heat_map_of_California) + geom_sf(aes(fill = Murder), color = "black", size = 0.5) + scale_fill_viridis_c(option = "mako" , begin = 0.1, label = scales::comma) + stat_sf_coordinates() + theme(legend.position="right",
        plot.title = element_text(hjust = 0.5,color = "dark red", size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,color = "dark red", size = 14, face = "bold"),
        plot.caption = element_text(color = "dark grey"))+ 
  guides(fill = guide_legend(title = "Murder Count", title.position = "top", title.theme =element_text(size = 10, face = "bold",colour = "Black",angle = 0))) + geom_sf_text(aes(label = NAME), colour = "dark grey", alpha = 0.6) + labs(title = "Murder Count by Cities in California", subtitle = "1980 to 2014")



```

In California most of their murder cases came from the city of Los Angeles and the southside of California compared to Northern California.

## 6. Genders of Victim by State

We count amount of cases based on data about genders by state. We took the dataset to combine the total amount of Female count by state with Male total count together into 1 dataset to compare it.


```{r, echo=FALSE, results=FALSE, fig.width= 18, fig.height= 15, message=FALSE, warning=FALSE}
Female_Data <-  data %>% group_by(State, Victim.Sex) %>% summarize(Female_Total = n()) %>% filter(Victim.Sex == "Female")
Male_Data <- data %>% group_by(State, Victim.Sex) %>% summarize(Male_Total = n()) %>% filter(Victim.Sex == "Male")
Unknown_Data <- data %>% group_by(State, Victim.Sex) %>% summarize(Unknown_Total = n()) %>% filter(Victim.Sex == "Unknown")
Gender_Data <- Female_Data %>% left_join(Male_Data, by = "State")
Gender_Data_with_Total<- Gender_Data %>% mutate(Sum_Gender = Female_Total + Male_Total)
Gender_Data_plus_Unknown <- Gender_Data %>% left_join(Unknown_Data, by = "State")
Gender_Data_for_3D <- Gender_Data_plus_Unknown %>% mutate(Unknown_Total = ifelse(is.na(Unknown_Total), 0, Unknown_Total)) %>% mutate(Sum_Gender = Female_Total + Male_Total + Unknown_Total) 
Gender_heatmap <- states %>% left_join(Gender_Data_for_3D, by = c("NAME" = "State"))
ggplot(Gender_heatmap) + 
  geom_sf(aes(fill = Female_Total), colour = alpha("black", 1 / 2), size = 0.2) +   
  scale_fill_viridis_c(option = "plasma" , 
                       begin = 0.1, name = "Murder Count",label = scales::comma) + geom_sf_text(aes(label = NAME), colour = "black") + 
  labs(title = "Heat Map by Female Murder Cases", 
      subtitle = " By each State") + theme(panel.background = element_rect(colour = "black")) + 
theme_bw() 
ggplot(Gender_heatmap) + 
  geom_sf(aes(fill = Male_Total), colour = alpha("black", 1 / 2), size = 0.2) +   
  scale_fill_viridis_c(option = "brewer blues" , 
                       begin = 0.1, name = "Murder Count",label = scales::comma) + geom_sf_text(aes(label = NAME), colour = "black") + 
  labs(title = "Heat Map by Male Murder Cases", 
      subtitle = "By each State") + theme(panel.background = element_rect(colour = "black")) + 
theme_bw() 
```

We made two heatmaps, one to show male murder cases and the other one to show the female murder cases in the US. We can see some similarities within male and female murders. They both have a lot of murder cases in California, Texas, and New York. 


## 7. Is there any missing or unknown data in this dataset?

```{r} 
#As an example, let's see how they show in gender of victims field!
data %>% group_by(Victim.Sex) %>% summarize(Gender = n())
```


In this table as you can see, we have 984 that are unknown so we need to tidy up our data and get rid of the unknowns.



```{r}

#How about Unknown Weapon?
data %>% group_by(Weapon) %>% 
  summarize(Most_Weapon_Used = n()) %>% 
  arrange(desc(Most_Weapon_Used)) %>% 
  kable() %>% kable_paper() %>% scroll_box(height = "300px")

```


You can see that handgun was 10x most used weapons compared to others. Knife was the second most used weapon. It is interesting know that the hand gun was the most used weapon, it makes us wonder if gun laws were regulated in the US would it make the handgun the least used of weapon and make the knife the most used weapon. And 33192 unknown weapons with many different kinds of guns need to be sorted out.



## Let's see how the distribution of cases by victims' ages is based on data we have!

```{r, message = F, fig.width=7, fig.height=5}
# Graph for cases by age
data %>% ggplot(aes(Victim.Age)) + geom_histogram(binwidth = 50) + 
  labs(title = "How many cases over victims' ages?", 
       x = "Age of Victim (years old)", y = "Cases")
```


From then, we see that there are many cases with nearly 1000 year-old victims. It doesn't make sense so then we proceeded to filter our data to make it more neat and coherent.



# Filtered Data  

```{r echo=FALSE, warning=FALSE}
filtered_data <- data %>% 
  select(Year, Victim.Age, Victim.Sex, Relationship, Weapon) %>% 
  filter(Relationship != 'Unknown') %>% 
  filter(Weapon != 'Unknown') %>% 
  filter(Victim.Sex != 'Unknown') %>% 
  filter(Victim.Age %in% c(1:100))
glimpse(filtered_data)

```

In our filtered data we decided to work with data that we would find useful for our findings and remove all the 'unknowns' in the dataset. In our new filtered data we decided to work with __Year, Victim's Age, Victim's Sex, Relationship to their perpetrator and type of Weapon__ used in each incident. We also filtered the victim's age to be more accurate from combining them from __age 1 to 100.__ 


# Analyze factors and link to our question

## 1. Year:

We wanted to observe the number of cases throughout the three decades. 


```{r, echo=FALSE, message=F,warning = F}
# How many cases by year?
filtered_data %>% count(Year)  %>% 
  kbl() %>% kable_paper()  %>% scroll_box(height = "300px")

```


Now with our filtered data, we wanted to see the crimes rates throughout the years of 1980-2014. In 1980 & 1993, you can see that there is a peak in crimes rates but then they start to decreased. In 1980, the crime rate was high due to a severe global economic recession and inflation peaked in the US by 14.76%


```{r,message=FALSE,fig.width=7, fig.height=5}
# The rate of murders during the period of 1980-2014
filtered_data %>% 
  group_by(Year) %>% 
  summarise(murder = n()) %>% 
  ggplot(aes(Year,murder)) + geom_point() + geom_smooth()
```



## 2. Relationship between them:

Then we wanted to see if the perpetrator had knew their victim before striking. So we made a variable based on their relationship. 

```{r,echo=FALSE, message=F, warning=F}
#Define if they know each other
relationship_data <- filtered_data %>% 
  mutate(Relationship_with_murder = (Relationship == "Stranger")) %>% 
  mutate_all(funs(replace(.,. == "TRUE", "Unknown"))) %>% 
  mutate_all (funs(replace(.,. == "FALSE", "Known")))
glimpse(relationship_data)

```

Looking at this table, there is a higher chance that the victim knows their perpetrator. It is a huge difference.

```{r}
#Count how many cases they know each other
relationship_data %>% group_by(Relationship_with_murder) %>% summarise(cases = n())
```


```{r,echo=F, message = F, warning = F, fig.width=7, fig.height=5}
# Graph for Relationship with Murder
table_relationship <- relationship_data %>% group_by(Relationship_with_murder) %>% summarise(cases = n())
blank_theme <- theme_minimal()+
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  plot.title=element_text(size=14, face="bold")
  )
table_relationship %>%
  ggplot(aes(x="", y= cases, fill = Relationship_with_murder)) + 
  geom_bar(color = 'white', width = 1, stat = 'identity') + 
  coord_polar("y", start=0) + 
  blank_theme + 
  theme(axis.text.x=element_blank()) +
  geom_text(aes(label = cases),
            position = position_stack(vjust = 0.5)) +
  labs(title = "Have they known each other?", fill = "Relationship with murder") 
```

In this pie chart, we wanted to know when the victim knew their perpetrator or had some type of relationship or if they were simply strangers. It is clear that the amount of victims who knew the murder is nearly three times than the amount of victims who were strangers to the perpetrators.


## 3. Gender:

We wanted to count the Victim Sex and see the graph so the data is filtered here and in my findings, there is a higher percent for a male to be murdered than a female.

```{r, echo = FALSE}
# I wanted to count the Victim Sex and see the graph 
filtered_data %>% count(Victim.Sex)
```

```{r,fig.width=7, fig.height=5}
# Graph for Victim Sex
filtered_data %>% ggplot(aes(Victim.Sex, fill = Victim.Sex)) + 
  geom_bar(color = 'black') + theme_bw() + 
  geom_text(aes(label = ..count..), stat = "count", vjust = 5) + 
  labs(title = "Which gender is the most targeted?", 
       x = "Victim Gender", y = "Cases", fill = "Victim Gender")


```

Since we made a table to show what gender is killed the most we wanted to see a bar chart to compare both genders. In our analysis, it helps to see how much more a male is to be murdered. 


Then we wanted to see the correlation based on the victim's gender and it's perpetrator relationship.

```{r,fig.width=7, fig.height=5}
# We wanted to see victim's gender and the correlation of their relationship?
relationship_data %>% ggplot(aes(Victim.Sex, fill = Relationship_with_murder)) + 
  geom_bar(color = 'black') + 
  theme_bw()+ 
  geom_text(aes(label = ..count..), stat = "count", vjust = 1) + 
  labs(title = "How many cases do they know each other by genders? ", 
       x = "Victim Gender", y = "Cases", fill = "Relationship with murder")

```

From above graph, it is obvious that most of victims know the murders before the incident



## 4. Age

__How does the distribution of cases look like by victim's age?__

```{r,fig.width=7, fig.height=5}
filtered_data %>% ggplot(aes(Victim.Age)) + 
  geom_histogram(color = 'Black', fill = 'white', binwidth = 3) +
  labs(x = "Victim Age", y = "Cases") 
```


_We could see that the average age of a victim to be most likely murdered are the ages 21-25. However, it is not really clear to determine if the age is an effected factor on the rate of murder cases. So, let's take a look at this flow._


```{r,echo=FALSE,message=F,warning=F}
library(gifski)
library(ggthemes)
table_age <- filtered_data %>% group_by(Victim.Age) %>% summarise(cases = n())
plot <- table_age %>%
  ggplot(aes(x=Victim.Age, y= cases)) + 
  geom_line(alpha=1,colour='white') + 
  theme_solarized_2(light=F)+
  labs(title="The change of murder cases over Victim's Ages", x = "Victim Age", y = "Cases") +
  theme(text=element_text(colour="#EEEEEE"),
        title=element_text(colour="#EEEEEE",size=9,face = "bold"),
        plot.title=element_text(hjust=0.5),
        axis.title.x = element_text(),
        panel.grid.minor.x=element_blank(),
        legend.background = element_blank(),
        legend.key= element_blank(),
        legend.position=c(0.095, 0.81), ## legend at top-left, inside the plot
        plot.margin = unit(c(0.5,1.3,0.5,0.5), "cm"))
plot.animation=plot+
  transition_reveal(Victim.Age)+
  view_follow(fixed_y=T)

 
animate(plot.animation, height=420,width=700,fps=30,duration=10,end_pause = 120,res=100,rewind=F, renderer = gifski_renderer())
anim_save("The change of murder cases over Victim's Ages.gif")
```




# Summary

To summarize everything some important takeaway from our analysis is that the perpetrator gets acquainted with the victims before committing murder. Men are more like to strike someone they know than a stranger. Women also are more likely to strike someone they know. This is important to know in today's world because we are more likely to be murdered by someone we know rather than by a stranger. We often find ourselves feeling more safe to be with someone we know rather than a stranger. Based on our findings, men are 3x more likely to be murdered than women.The majority of homicide victims are around 30 years old. However, this age is not exclusive, and victims can be of any age. Also, the perpetrator usually is between the ages of 21 and 25. In comparison, the victim profile is more common in older age brackets. Lastly, the most used weapon used in a crime scene has been a handgun. It makes us wonder if gun law's were regulated in each state, would it reduce crime since it is easily accessible to acquire one. The underlying factors and motives of a serial killer is that they all may have different motives, where it can be in desperate need of money, power, sex, etc but something about them is that there are always to strike again.
